---
title: "Data_Analysis"
author: "Caroline Mayberry"
date: "`r Sys.Date()`"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set(message = FALSE) #whether to display messages
knitr::opts_chunk$set(collapse = TRUE) #less spacing between source code and text output blocks when knit

#make sure script produces an output file if testing:
knitr::opts_chunk$set(error = TRUE) #keep compiling output document even if there is an error
```

# [Data Wrangling]{.ul}

```{r packages}

#Read/export files
library(readxl) #excel
library(haven) #SPSS

#Format data frames
library(dplyr)
library(tidyverse)
library(tidyr)
library(reshape2)

#Analyze
library(PerformanceAnalytics)
library(corrplot)
library(Hmisc)
library(expss) #cross tabulation
library(lme4) #MLM
library(performance) #test for multicollinearity
library(effects) #glmer effects and summary tables
#library(glmmTMB) #binomial models with overdispersion
library(lmerTest) #prints p-values
options(scipen=999) #turns off scientific notation
options(digits=3) #number of significant digits when printing values

#Plot
library(ggplot2)
library(classifierplots)
library(yardstick)
library(RColorBrewer)
library(dichromat)
library(ggeffects) #glmm marginal effects
library(jtools) #plot glmm effects and predicted effects with confidence intervals

```

**Reference Guides for R Packages:**

* [ggpredict vs. ggeffects](https://strengejacke.github.io/ggeffects/reference/ggpredict.html)
  + ggpredict() uses the reference level, while ggeffect() and ggemmeans() compute a kind of "average" value, which represents the proportions of each factor's category.  
* [ggeffects (for random effects)](https://cran.r-project.org/web/packages/ggeffects/vignettes/introduction_randomeffects.html)
  + the fixed effects and zero-inflation component (population-level), taking the random-effect variances into account (type = "zi_random"). Similar to mixed models without zero-inflation component, type = "fixed" and type = "random" for glmmTMB-models (with zero-inflation) both return predictions on the population-level, where the latter option accounts for the uncertainty of the random effects.
* [performance](https://easystats.github.io/see/articles/performance.html) 
  + test for multicollinearity

## Tidy Data

Recode levels of a factor variable:
```{r Recode lvls factor var}
levels(data$var.f)[levels(data$var.f) == "x"] <- "y"
```

Centering:
```{r Centering}
#Center Age
AllData$Age.cent <- AllData$Age - mean(AllData$Age)
```

Group Mean Centering:
```{r Group Mean Centering}
#Mean Center Pen Lifts for each Letter group
AllData <- AllData %>%
  group_by(Letter) %>%
  mutate(PenLifts.letterMean = mean(PenLifts.Freq)) #mean PenLifts for letter

AllData <- AllData %>%
  mutate(PenLifts.c = (PenLifts.Freq - PenLifts.letterMean)) #center scores by letter group mean

plot(AllData$PenLifts.Freq)
plot(AllData$PenLifts.c)
```

Check Missing Data:
```{r Check Missing Data}
cro(df$ID, df$var)
```


Subset data:
```{r Subset}
#by column
corData <- subset(wideData, select = c(Age, var1, var2))

#by value
subset <- data[which(data$var == "y"),]
```

Dummy code:
```{r Dummy code}
data$DummyCode <- NA
data$DummyCode[which(data$var == "n")] <- 0
data$DummyCode[which(data$var == "y")] <- 1
```


Match levels of var in two files/data frames: 
```{r Merge by factor lvl}
#set ID as factor
data1$ID <- as.factor(data1$ID)
data2$ID <- as.factor(data2$ID)

#filter for only IDs that appear in data2
data1filtered <- data1[data1$ID %in% data2$ID,]

#remove factor levels for IDs that were filtered out
data1filtered$ID <- droplevels(data1filtered$ID)
```

Add to repeated measures data set (loop through all rows):
```{r fill RM data, echo = FALSE, warning = FALSE}
rmData$Age <- NA
rmData$Sex <- NA

for (i in levels(rmData$ID)){
  rmData$Age[rmData$ID == i] <- wideData$Age[wideData$ID == i]
  rmData$Sex[rmData$ID == i] <- wideData$Sex[wideData$ID == i]
}
```

Calculate count/frequency from repeated measures and add to wide format data:
```{r Get count from RM}
wideData$Freq <- NA

#loop through each subject
for (id in levels(rmData$ID)){
  #set count to 0 each time
  count <- 0
  #subset by participant
  idSet <- subset(rmData, rmData$ID == id)
  #loop rows associated w that ID# and count occurence of var
  for (r in 1:nrow(idSet)){
    if (idSet$var[r] == "y"){
      count <- count + 1
    }
  }
  #input count for each subj
  wideData$Freq[wideData$ID == id] <- count
}
```

------------------------------------------------------------------------

# [Desc Stats, Freq & Plots]{.ul}

```{r Packages}
library(ggplot2)
```

Print descriptive statistics for subjects:
```{r Sample Descriptives, echo = FALSE}
#N and n female
cat("N = ", length(data$ID), ", (", length(subset(data$Sex, data$Sex == "F")), "F )", "\n")

#mean age 
cat("mean age (mo): ", mean(data$AgeMo),"\n")
#SD age
cat("sd age (mo): ", sd(data$AgeMo),"\n")
#Age range
cat("range: " , range(subjData$Age.in.Months))

```

<!-- # library(skimr) -->
<!-- # DescStats <- skim(SubjTotalScores) -->
<!-- #  -->
<!-- # DescStats.pK <- skim(SubjTotalScoresPK[which(SubjTotalScoresPK$Kgrd == 0),]) -->
<!-- # DescStats.K <- skim(SubjTotalScoresPK[which(SubjTotalScoresPK$Kgrd == 1),]) -->


Frequency Tables:

[Frequencies and Crosstabs Tutorial](https://www.statmethods.net/stats/frequencies.html)

```{r Frequency Tables}
library(expss)

#one var
cat("Overall event frequencies:")
cro(dvData$Event)
cat("\n\n")

#crosstab
cat("Events by Age(months):")
cro(dvData$Event.f, dvData$Age)
cat("\n\n")
```


Histograms:
```{r Histograms}
#For all columns of a dataframe:
hist(AvgSubjData)

#Plot each subject, sorted by age, segmented bar for factor var
ggplot(data, aes(x=reorder(ID, Age), fill = factor.var)) +
   geom_bar()

#Same, but with facet_wrap to separate plots by second factor var
ggplot(data, aes(x=reorder(ID, Age), fill = factor.var)) +
   geom_bar()+
   facet_wrap(as.factor(data$factor.var2))
```

```{r histograms loop}
#select variables
 distData <- subset(allData, select = c(var1, var2))

#Change columns to numeric type
distData <- distData %>% mutate_if(is.character,as.numeric)
distData <- distData %>% mutate_if(is.factor,as.numeric)
for( i in 1:ncol(distData)){
  hist(distData[,i], main = colnames(distData)[i],xlab = colnames(distData)[i], col = 'pink')
}



#break down by group
distDataPK <- distData[which(distData$GradeGroup == "0"), ]
for( i in 1:ncol(distDataPK)){
  hist(distDataPK[,i], main = colnames(distDataPK)[i],xlab = colnames(distDataPK)[i], col = 'light blue')
}

distDataK <- distData[which(distData$GradeGroup == "1"), ]
for( i in 1:ncol(distDataK)){
  hist(distDataK[,i], main = colnames(distDataK)[i],xlab = colnames(distDataK)[i], col = 'blue')
}

#Check dist of each variable
variableData <- subset(binaryData, select = c("var1", "var2"))
for( i in 1:ncol(variableData)){
  hist(variableData[,i], main = colnames(variableData)[i],xlab = colnames(variableData)[i], col = 'yellow')
}
```


Scatter Plot:
```{r Scatter Plot}
#basic
plot(wideData$Age, wideData$var, main="Age x Var",
   xlab="Age (mo) ", ylab="Var", pch=19)

#ggplot2
ggplot(data = data, aes(x= var1, y = var2)) +
  geom_point() +
  labs(title = "Title") +
  theme_minimal()

#ggplot2 with color grouped points
ggplot(data = data, aes(x= var1, y = var2, colour = factor.var)) +
  geom_point() +
  labs(title = "Title") +
  theme_minimal()
```

Counts Plot:
- to overcome the problem of data points overlap is to use what is called a counts chart. Whereever there is more points overlap, the size of the circle gets bigger.
[Counts Plot Source](http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html#Counts%20Chart)
```{r Counts Plot}
theme_set(theme_bw())  # pre-set the bw theme.
g <- ggplot(mpg, aes(cty, hwy))
g + geom_count(col="tomato3", show.legend=F) +
  labs(subtitle="mpg: city vs highway mileage", 
       y="hwy", 
       x="cty", 
       title="Counts Plot")
```


Boxplot:
```{r Boxplot}
freq_table <- data.frame(table(data$ID, data$var.f))

ggplot(freq_table, aes(x = var.f, y = Freq)) + 
  geom_boxplot() + 
  ylab(names(levels(freq_table$var.f)))

```

Marginal Histogram/Boxplot:
[Marginal Plot Source](http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html#Counts%20Chart)
```{r Marginal Plots}
# load package and data
library(ggplot2)
library(ggExtra)
data(mpg, package="ggplot2")
# mpg <- read.csv("http://goo.gl/uEeRGu")

# Scatterplot
theme_set(theme_bw())  # pre-set the bw theme.
mpg_select <- mpg[mpg$hwy >= 35 & mpg$cty > 27, ]
g <- ggplot(mpg, aes(cty, hwy)) + 
  geom_count() + 
  geom_smooth(method="lm", se=F)

ggMarginal(g, type = "histogram", fill="transparent")
ggMarginal(g, type = "boxplot", fill="transparent")
# ggMarginal(g, type = "density", fill="transparent")
```


------------------------------------------------------------------------

# [ggplot2]{.ul}

[Cheat Sheet](https://www.maths.usyd.edu.au/u/UG/SM/STAT3022/r/current/Misc/data-visualization-2.1.pdf)

### Themes

```{r ggplot themes}
p + theme_grey() #default
p + theme_bw()
p + theme_minimal() 
p + theme_classic() 
```

### Labels

```{r labelled ggplot}
ggplot(data = mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  labs(title = "Fuel Efficiency by Car Weight",
       subtitle = "Motor Trend Magazine 1973",
       caption = "source: mtcars dataset",
       x = "Weight (1000 lbs)",
       y = "Miles per gallon") +
 
   theme(plot.title    = element_text(family = "bell", size=14),
         plot.subtitle = element_text(family = "gochi"),
         plot.caption  = element_text(family = "caveat", size=15),
         axis.title    = element_text(family = "comic"),
         axis.text     = element_text(family = "comic", 
                                      face="italic", size=8))
```

------------------------------------------------------------------------

# [Correlations]{.ul}

```{r Correlation Matrix}
#subset only columns that will be in correlation analysis
corData <- subset(wideData, select = c(Age, var1, var2))

library(PerformanceAnalytics)
chart.Correlation(corData, histogram=FALSE, pch = "+")

library(Hmisc)
variablesCor <- rcorr(as.matrix(corData))
corrplot(variablesCor$r, type="upper", 
         p.mat = variablesCor$P, 
         sig.level = 0.05, insig = "blank", tl.col="black", 
         tl.srt=45, diag=FALSE, addCoef.col = "black")
```

## Co-occurence of types of handwriting errors
Look at co-occurence of types of handwriting errors
(Source)[https://towardsdatascience.com/how-to-create-co-occurrence-networks-with-the-r-packages-cooccur-and-visnetwork-f6e1ceb1c523]
```{r Co-occurence, layout="l-body",  echo=FALSE}
# #Transpose data
# dataTransposed <- subset(errorCoded, select = -c(ID, variable, Visit, HandwritingSum, HandwritingBinary))
# dataTransposed[is.na(dataTransposed)] <- 0
# dataTransposed <- t(dataTransposed)
# 
# library(cooccur)
# 
# #Print significant pairwise co-occurrences and their labels.
# co <- print(cooccur(dataTransposed, spp_names = TRUE))
# 
# detach("package:cooccur", unload = TRUE)
```


# [Modeling]{.ul}

### Family Objects for Models

> | binomial(link = "logit")
> | gaussian(link = "identity")
> | Gamma(link = "inverse")
> | inverse.gaussian(link = "1/mu\^2")
> | poisson(link = "log")
> | quasi(link = "identity", variance = "constant")
> | quasibinomial(link = "logit")
> | quasipoisson(link = "log")


### Save to .doc Format
```{r model .doc output table}
#Save output table to a .doc file
tab_model(uncond_model, randomslope_model, show.re.var = TRUE, show.stat = TRUE, show.se=TRUE, show.ci = FALSE, show.reflvl = TRUE,  p.style = "stars", file = "final model.doc")
```


### Simple Linear Models

```{r lm}
#Usage lm {stats}
lm(formula, data, subset, weights, na.action,
   method = "qr", model = TRUE, x = FALSE, y = FALSE, qr = TRUE,
   singular.ok = TRUE, contrasts = NULL, offset, ...)
```

#### Plot

```{r plot points}
ggplot(subjData, aes(x, y)) +
  geom_point() +
  stat_smooth(method = lm)
```

### GLM

Generalized Linear Models

```{r GLM}
#Usage {stats}
glm(formula, family = gaussian, data, weights, subset,
    na.action, start = NULL, etastart, mustart, offset,
    control = list(...), model = TRUE, method = "glm.fit",
    x = FALSE, y = TRUE, singular.ok = TRUE, contrasts = NULL, ...)

glm.fit(x, y, weights = rep.int(1, nobs),
        start = NULL, etastart = NULL, mustart = NULL,
        offset = rep.int(0, nobs), family = gaussian(),
        control = list(), intercept = TRUE, singular.ok = TRUE)

#example w normal dist
summary(m1 <- glm(var1 ~ Age, family="gaussian", data=wideData))
```

##### Plot effects

```{r effects plot}
library(effects)
plot(allEffects(m1))
```

### Linear Mixed-Effects Models

> The main features distinguishing lme4 from nlme are:
>
> 1.  more efficient linear algebra tools, giving improved performance on large problems
> 2.  simpler syntax and more efficient implementation for fitting models with crossed random effects
> 3.  the implementation of profile likelihood confidence intervals on random-effects parameters
> 4.  the ability to fit generalized linear mixed models
>
> @bates2015

```{r LME}
library(lme4)
summary(m <- glmer(Success ~ Age.c + (1|ID),
                   family="poisson", 
                   data = rmData))

library(effects)
plot(allEffects(m))
```

#### Binary Logistic Regression (with crossed random effects)

```{r Cross-Classified Binary Logistic Regression}
#binary logistic regression
fullMod <- glmer(y ~ x1 + x2 + (1|ID) + (1|Group),
              family=binomial("logit"), 
              data=binaryData)
summary(fullMod)

ggpredict(fullModLBb, "x1")
ggpredict(fullModLBb, "x2")

library("effects")
e1.lm1 <- predictorEffect("x1", fullMod)
plot(e1.lm1)
```

### Zero-Inflated Poisson Regression

# [Inter-Rater Reliability]{.ul}

```{r IRR}
#subset column of scores from each rater
IRRdata <- cbind(data_c1$var, data_c2$var)

#percent agreement
agree(IRRdata)

#Unweighted Kappa 
kappa2(IRRdata, "unweighted")

#ICC 
icc(IRRdata, model = "twoway", type = "consistency", unit = "single")
icc(IRRdata, model = "twoway", type = "consistency", unit = "average")
icc(IRRdata, model = "twoway", type = "agreement", unit = "single")
icc(IRRdata, model = "twoway", type = "agreement", unit = "average")
```
